---
title: 'Part 1:  exponential distribution'
author: "Swathi Vijayakumar"
date: "August 23, 2015"
output: html_document
---
In purpose of part 1 is to investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution will be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. We will set lambda = 0.2 for all of the simulations and investigate the distribution of averages of 40 exponentials. At least a thousand simulations will be performed. 

###1. Show the sample mean and compare it to the theoretical mean of the distribution.
```{r}
library(stats)
n = 40
nosim = 1:1000
set.seed(5)
lam = 0.2
dat = data.frame(exp_dist = sapply(nosim,function(x){mean(rexp(n,lam))}))
```


**first calculate the mean of the exponential distribution**

```{r}
samp_mean = mean(dat$exp_dist)
round(samp_mean,3) 
```

we are told that the mean of the exponential distribution is 1/lambda

```{r}
round(1/lam,3)
```
As shown above the actual center of the distribution 5.043 is close to the thoretical center of the distribution 5.


###2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.


**Calculate the standard deviation of the exponential distribution using the sd function**

```{r}
sd = sd(dat$exp_dist)
round(sd,3)
```
The theoretical standard deviation is
```{r}
round((1/lam)/sqrt(40),3)
```

**Calculate the variance of the exponential distribution using the var function**

``` {r}
samp_var = var(dat$exp_dist)
round(samp_var,3)
```

we are told that the standard deviation is 1/lambda. The theoretical Variance is sigma^2/n where sigma is the standard deviation

```{r}
round(((1/lam)^2)/n,3)
```

As shown above the actual variance 0.603 is close to the thoretical variance of 0.625.

###3. Show that the distribution is approximately normal.

```{r}
library(ggplot2)
g = ggplot(dat, aes(x=exp_dist)) + geom_histogram(binwidth=.3,color = "black",fill = "darkred",aes(y=..density..)) 
g = g + stat_function(fun=dnorm,arg = list(mean = 5,sd = sd(dat$exp_dist)),size=2, warning=FALSE) 
g
````

The histogram plot above shows that the distribution is approximately normal (shown by curve in black) with a mean of 5 and standard deviation of 0.791